Efficient Handling of String-Number Conversion

### 2019-11-21 draft.tex
formulas to formulae globally

Our decision procedure has two steps: the first step consists in over-approximating the set of input constraints into a set that falls in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Observe that we could  over-approximate the input constraint into any decidable fragment, e.g. the acyclic fragment~\cite{abdulla2014string} or the straight-line fragment~\cite{chen2019decision}. Our choice of the  chain-free fragment~\cite{abdulla2019chain} is only motivated by the fact that the chain-free fragment is the {\em largest} known decidable fragment for that class of string constraints. In our example, we over-approximate the formula $\Phi$ by converting $xy=yx$ to two formulas $\{x_1=xy, x_2=yx\}$  and replacing the constraint $n=\sti{x}$ with $n=-1 \vee (n\neq -1 \wedge x\in [0-9]^*$). The over-approximation module uses also some heuristics to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume in some case, we could derive that the value of $n=12$ from some integer constraints.  Then, we could further derive the value of $x$ is in the regular language $(0^*12)$. 
Observe that if the over-approximation is UNSAT then our decision procedure  declares that the original formula is also UNSAT and terminates. Surprisingly, despite its simplicity, our over-approximation procedure works very well in practice. Coming back to the formula $\Phi$, the over-approximation module will  return SAT  in this case.

@@

%The second step of our decision procedure is only enabled if the over-approximation step returns SAT. In this case, our decision procedure restricts the domain of each string variable to strings that obey some predefined and parameterized pattern.  Adjusting the parameters enlarges or prunes the potential solution space. In this paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA). A PFA is a {\em flat} finite state automaton consisting of a predefined sequence of loops, each of fixed length (see Figure \ref{fig:sfa_def}). The approach based on PFA is very flexible yet allows very efficient manipulation. One can pick PFA of \emph{arbitrary structure} for search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Furthermore, we associate a unique natural number with each character. This allows us to assume that the input alphabet is a subset of natural numbers to avoid the \textit{alphabet explosion problem} from which the approach in~\cite{abdulla2017flatten} suffers.  We label each transition between two states of a PFA with a unique \emph{character} variable (whose domain is the set of natural numbers) instead of having a transition between each two states for each symbol in the alphabet.  
%Such reduction based on PFA is the key to handling string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to the satisfiability problem of linear formulas and can be solved efficiently.

%The battle field is in fact those SAT instances, for which we develop solutions using under-approximation. For under-approximation, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA) (will be defined formally in Section~\ref{section:sfa}). The approach based on PFA is very flexible yet allows very efficiently manipulation. One can pick PFA of \emph{arbitrary structure} for the search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Moreover, it avoids the \textit{alphabet explosion problem} that the approach in~\cite{abdulla2017flatten} suffers. 
%%In contrast to under-approximation based on bit-blasting~\cite{kiezun2009hampi} (i.e., limit the length of the strings and encode the string constraint to a Boolean formula), our linear formula for under-approximation still encodes possibility an infinite number of models. 
%Such reduction based on PFA is the key to handle string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to linear formulas and can be solved efficiently. We can step-wisely enrich the PFAs and hence solve the satisfiability problem of string constraints in a systematic manner.

@@

%The battle field is in fact those SAT instances, for which we develop solutions using under-approximation. For under-approximation, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA) (will be defined formally in Section~\ref{section:sfa}). The approach based on PFA is very flexible yet allows very efficiently manipulation. One can pick PFA of \emph{arbitrary structure} for the search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Moreover, it avoids the \textit{alphabet explosion problem} that the approach in~\cite{abdulla2017flatten} suffers. 
%%In contrast to under-approximation based on bit-blasting~\cite{kiezun2009hampi} (i.e., limit the length of the strings and encode the string constraint to a Boolean formula), our linear formula for under-approximation still encodes possibility an infinite number of models. 
%Such reduction based on PFA is the key to handle string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to linear formulas and can be solved efficiently. We can step-wisely enrich the PFAs and hence solve the satisfiability problem of string constraints in a systematic manner.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Overview} \label{section:overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%Here we try to explain the approach using a toy example $\Phi = \{xy=yx, n = \sti{x}, n >3 ,|y|>|x|, y \in (12)^+\}$. Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12"$ and $y=``1212"$.
%
%For proving UNSAT, our framework does not set any restriction on which decision procedure to use. One can select any decision procedure they want to use and ``over-approximate" the input constraint to one that falls in the decidable fragment, e.g., the acyclic fragment~\cite{abdulla2014string}, the straight-line fragment~\cite{chen2019decision}, or the chain-free fragment~\cite{abdulla2019chain}. For instance, we can convert $xy=yx$ to two formulas $\{x_1=xy, x_2=yx\}$ using two auxiliary variables $x_1,x_2$ and replace the constraint $n=\sti{x}$ with $n=-1 \vee (n\neq -1 \wedge x\in [0-9]^*$). This produce new formulas that over-approximate the original ones and fall in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Since $\Phi$ is satisfiable, the over-approximation module will not return UNSAT. In our implementation \tool, we further add heuristics trying to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume in some other cases, we can derive the value of $n=12$ from some integer constraints. In such case, we can further derive the value of $x$ is in the regular language $(0^*12)$. 
%
%Then our procedure will restrict the domain of each string variables to the language defined by some parametric flat automata (PFA) and convert them to equisatisfiable linear formula. The details of the under-approximation component will be described in Section~\ref{section:under_approximate}. Here we try to explain what the linear formula looks like by an example. Assume that we project the domain of $x$ and $y$ to the PFA in Figure~\ref{fig:sfa} (a) and (b), respectively. The variables $v_1$, $v_2$, $v_3$, $v_4$ in the figure are \emph{character} variables. One special treatment we do in this paper is that we encoding characters as natural numbers. So $v_1$, $v_2$, $v_3$, $v_4$ are also integer variables. For example, from the formula $ y \in (12)^+$, we may derive $v_3=1$ and $v_4=2$. 
%
%The linear formula produced after the domain restriction will be over variables $v_1$, $v_2$, $v_3$, $v_4$, as well as, the number of occurrences of each character variables $\#v_1$, $\#v_2$, $\#v_3$, $\#v_4$. Each model of the linear formula encodes a model of the string constraint. For example, $x=``12"$ and $y=``1212"$ is encoded by the assignment $(v_1,v_2,v_3,v_4,\#v_1,\#v_2,\#v_3,\#v_4)= (1,2,1,2,1,1,2,2)$. The assignment says, for example, that $x$ is the parametric word obtained by traversing the loop of $A_x$ once (because $\#v_1 = \#v_2 = 1$), which is $v_1v_2$. Under the assignment $v_1=1$ and $v_2=2$, we obtain $x=``12"$.
%
%If a model of the produced linear formula is found, then the procedure concludes SAT with an assignment to the string variables. If not, our procedure step-wisely change the PFA to more expressive ones (by adding more states and transitions) to systematically search for the solution. Depending on the implementation, one may choose to report unknown after some certain steps.
%

@@

%Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.  
%Given a linear formula $\phi$, a set of variables $V$, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
%we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$. 
%
%\lh{Our use of indexing by $V$ in the text later is not very nice. Add this for clarity?: If $V$ contain additional variables not present in $\phi$, the these might be assigned arbitrary values by $I$. If $V$ is missing some of the variables of $\phi$, then it must be possible to complete $I$ with assignment of the missing variables into a satisfying assignment of $\phi$ (as in existential quantification).}
%We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
%We drop the reference to $V$ and write just ``model of $\phi$", $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.  
Given a linear formula $\phi$ over variables $V$ and an \emph{integer interpretation} of $V$, a function $I:V\rightarrow \integers$, 
we denote by $I\models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$. 
We use $\modelsof \phi$ to denote the set all models of $\phi$. 
%
%Let $J$ and $J'$ be two sets integer interpretations, $J$ containting interpretation of variables $V$ and $J'$ of variables $V'$. 
%We say that $J$ and $J'$ are \emph{equivalent}, written $J\iequiv J'$, if their interpretations restricted to the common variables are the same, that is, if 
%$\{\restrict I {V\cap V'} \mid I \in J\} = \{\restrict {I'} {V\cap V'} \mid I' \in J'\}$. We may write $J\eqwrt{V\cap V'} J'$ if we wish to made the common variables explicit.

@@

%\section{Under-approximation} \label{section:under_approximate}
%In this section, we describe formally how to restrict the domain of string variables to patterns defined by \emph{parametric flat automata} (PFA) (Section~\ref{section:sfa}) and hence convert string constraints $\insc$ to a set of equisatisfiable linear formulas $\outlf$ in polynomial-time. We assume all disequality constraint $t_1 \neq t2$ are already converted to equal-satisfiable equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}. 
%
%We pick a PFA for each variable in $\vars$ at the beginning of the decision procedure and then construct the set of linear integer constraints in a stepwise manner. In each step, we take a string constraint from $\insc$ and depending on the type of the string constraint, we execute the corresponding procedure to produce the linear integer constraints and add them to the output linear formulas $\outlf$. We also add constraints to $\outlf$ to ensure all character variables of the PFAs only pick values from $\Sigma_\epsilon$.
%
%More concretely, at the beginning, we move all integer constraints in $\insc$ to the output linear formulas $\outlf$. After this step, $\insc$ is the union of the three sets of constraints: the equality constraints, the regular constraints, and the string-number conversion constraints. For all string variables $x$ that occurs in a string-number conversion constraint $n= \sti{x}$, we follow the rule that will be described in Section~\ref{section:s2i} to pick the corresponding PFA. We do not set any restriction to the selection of PFA for other string variables. The selection is an implementation decision. We will describe the PFA selection strategy of \tool in Section~\ref{section:implementation}. 
%
%After the PFA for each variable is selected, we step-wisely pick and remove a string constraint from $\insc$, executing the corresponding procedure to create equisatisfiable linear constraints, and add all the results to $\outlf$. The procedures for regular, equality and string-number conversion constraints are in Sections~\ref{section:mem},~\ref{section:eq}, and~\ref{section:s2i}, respectively. In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.
%
%
% 
%
%
%
%\hide{We replace it with the following constraints.
%	
%	$$
%	\begin{array}{c}
%	(|x_1|+ |x_2|+ \cdots +|x_n| \neq |x_{n+1}| + |x_{n+2}|+ \cdots +|x_m|)\vee \\
%	\left(
%	\begin{array}{cccc}
%	
%	y_1\cdot y_2\cdot y_3 &=& x_1\cdot x_2 \cdots x_n &\wedge\\
%	y_1 \cdot y'_2 \cdot y'_3 &=& x_{n+1}\cdot x_{n+2} \cdots x_m &\wedge\\
%	|y_2|&=&|y'_2|=1
%	\end{array}
%	
%	\right)
%	\end{array}
%	$$
%	
%	The second disjunct says that the constraints $x_1\cdot x_2 \cdots x_n$ and $x_{n+1}\cdot x_{n+2} \cdots x_m$ have a common prefix $y_1$, but the next character $y_2$ and $y'_2$ is different. In an more efficient implementation, one just project $y_2$ and $y'_2$ to a PFA that accept only symbolic words $v_1$ and $v'_1$, respectively (the PFA has no loop and only one transition). }
%
%
%\yfc{say handling regular constraint can be exponential to the number of constraints, but in our under-approximation, it is polynomial.}

@@

\paragraph{Encoding epsilon.}
A minor technical difficulty is that we wish to build flat under-approximation in the form of pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, but $\epsilon$ is not a number. 
Hence, to allow purely arithmetic formulas talk about $\epsilon$, 
we encode $\epsilon$ as some fixed number $\noepsilon\in\nat\setminus\Sigma$.
%In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

%$$ % , to the output linear formulas $\phi_O$. 

@@

%Here we give a concrete example. For the equality constraint $xy = yx$, assume that we restrict the domains of $x$ and $y$ to the PFA in Figure~\ref{fig:sfa} (a) and (b), respectively. In the construction of the PFA for $xy$ and $yx$ (Figure~\ref{fig:sfa} (c) and (d)), we rename the variables on the PFA of $X$ and $Y$ to some fresh variable for each different occurrence of $x$ and $y$ in $xy = yx$, connect the PFA after renaming with ``epsilon" transitions (those labeled $v_{\epsilon_0}$ and $v_{\epsilon_1}$), and add linear formulas $v_{\epsilon_0}=-1$ and $v_{\epsilon_1}=-1$ to the output linear formulas $\phi_O$.

@@

%Recall that for flat automata, the number of occurrences of each symbol uniquely characterizes an accepting string (Section~\ref{section:sfa}) and for any finite automaton we can construct a Presburger formula characterizing its Parikh image (Section~\ref{section:preliminary}). For a product automaton $A$ computed from the previous step, we compute a Presburger formula $\phi(A)$ over the set of variables $\{\#{(v_i^{(j)},v_{i'}^{(j')})}\mid (v_i^{(j)},v_{i'}^{(j')}) \in \cvarone \times \cvartwo\}$ characterizing the Parikh image of $A$. We use $\#(v_i)$ and $\#(v_i^{(j)})$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively. The following formulas establish the relation of the number of occurrences of symbols between the PFAs and the product automaton.

>>>>

Our decision procedure has two steps: the first step consists in over-approximating the set of input constraints into a set that falls in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Observe that we could  over-approximate the input constraint into any decidable fragment, e.g. the acyclic fragment~\cite{abdulla2014string} or the straight-line fragment~\cite{chen2019decision}. Our choice of the  chain-free fragment~\cite{abdulla2019chain} is only motivated by the fact that the chain-free fragment is the {\em largest} known decidable fragment for that class of string constraints. In our example, we over-approximate the formula $\Phi$ by converting $xy=yx$ to two formulae $\{x_1=xy, x_2=yx\}$  and replacing the constraint $n=\sti{x}$ with $n=-1 \vee (n\neq -1 \wedge x\in [0-9]^*$). The over-approximation module uses also some heuristics to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume in some case, we could derive that the value of $n=12$ from some integer constraints.  Then, we could further derive the value of $x$ is in the regular language $(0^*12)$. 
Observe that if the over-approximation is UNSAT then our decision procedure  declares that the original formula is also UNSAT and terminates. Surprisingly, despite its simplicity, our over-approximation procedure works very well in practice. Coming back to the formula $\Phi$, the over-approximation module will  return SAT  in this case.

@@

%The second step of our decision procedure is only enabled if the over-approximation step returns SAT. In this case, our decision procedure restricts the domain of each string variable to strings that obey some predefined and parameterized pattern.  Adjusting the parameters enlarges or prunes the potential solution space. In this paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA). A PFA is a {\em flat} finite state automaton consisting of a predefined sequence of loops, each of fixed length (see Figure \ref{fig:sfa_def}). The approach based on PFA is very flexible yet allows very efficient manipulation. One can pick PFA of \emph{arbitrary structure} for search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Furthermore, we associate a unique natural number with each character. This allows us to assume that the input alphabet is a subset of natural numbers to avoid the \textit{alphabet explosion problem} from which the approach in~\cite{abdulla2017flatten} suffers.  We label each transition between two states of a PFA with a unique \emph{character} variable (whose domain is the set of natural numbers) instead of having a transition between each two states for each symbol in the alphabet.  
%Such reduction based on PFA is the key to handling string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to the satisfiability problem of linear formulae and can be solved efficiently.

%The battle field is in fact those SAT instances, for which we develop solutions using under-approximation. For under-approximation, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA) (will be defined formally in Section~\ref{section:sfa}). The approach based on PFA is very flexible yet allows very efficiently manipulation. One can pick PFA of \emph{arbitrary structure} for the search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Moreover, it avoids the \textit{alphabet explosion problem} that the approach in~\cite{abdulla2017flatten} suffers. 
%%In contrast to under-approximation based on bit-blasting~\cite{kiezun2009hampi} (i.e., limit the length of the strings and encode the string constraint to a Boolean formula), our linear formula for under-approximation still encodes possibility an infinite number of models. 
%Such reduction based on PFA is the key to handle string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to linear formulae and can be solved efficiently. We can step-wisely enrich the PFAs and hence solve the satisfiability problem of string constraints in a systematic manner.

@@

%The battle field is in fact those SAT instances, for which we develop solutions using under-approximation. For under-approximation, we restrict the search space of each string variable to strings that obey some predefined and parameterized pattern. By adjusting the parameters, one can easily enlarge or prune out the potential solution space. In paper, we propose to use patterns defined by \emph{parametric flat automata} (PFA) (will be defined formally in Section~\ref{section:sfa}). The approach based on PFA is very flexible yet allows very efficiently manipulation. One can pick PFA of \emph{arbitrary structure} for the search space restriction and reduce the string constraint solving problem to a linear formula satisfiability problem in polynomial-time. Moreover, it avoids the \textit{alphabet explosion problem} that the approach in~\cite{abdulla2017flatten} suffers. 
%%In contrast to under-approximation based on bit-blasting~\cite{kiezun2009hampi} (i.e., limit the length of the strings and encode the string constraint to a Boolean formula), our linear formula for under-approximation still encodes possibility an infinite number of models. 
%Such reduction based on PFA is the key to handle string-number conversion efficiently. When we restrict the solution space to those defined by PFAs, the original undecidable problem (i.e., string constraint with string-number conversion) is reduced to linear formulae and can be solved efficiently. We can step-wisely enrich the PFAs and hence solve the satisfiability problem of string constraints in a systematic manner.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Overview} \label{section:overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%Here we try to explain the approach using a toy example $\Phi = \{xy=yx, n = \sti{x}, n >3 ,|y|>|x|, y \in (12)^+\}$. Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12"$ and $y=``1212"$.
%
%For proving UNSAT, our framework does not set any restriction on which decision procedure to use. One can select any decision procedure they want to use and ``over-approximate" the input constraint to one that falls in the decidable fragment, e.g., the acyclic fragment~\cite{abdulla2014string}, the straight-line fragment~\cite{chen2019decision}, or the chain-free fragment~\cite{abdulla2019chain}. For instance, we can convert $xy=yx$ to two formulae $\{x_1=xy, x_2=yx\}$ using two auxiliary variables $x_1,x_2$ and replace the constraint $n=\sti{x}$ with $n=-1 \vee (n\neq -1 \wedge x\in [0-9]^*$). This produce new formulae that over-approximate the original ones and fall in the chain-free fragment~\cite{abdulla2019chain}, which is decidable. Since $\Phi$ is satisfiable, the over-approximation module will not return UNSAT. In our implementation \tool, we further add heuristics trying to derive the constant value of any side of the constraint $n=\sti{x}$ to refine the over-approximation. For instance, assume in some other cases, we can derive the value of $n=12$ from some integer constraints. In such case, we can further derive the value of $x$ is in the regular language $(0^*12)$. 
%
%Then our procedure will restrict the domain of each string variables to the language defined by some parametric flat automata (PFA) and convert them to equisatisfiable linear formula. The details of the under-approximation component will be described in Section~\ref{section:under_approximate}. Here we try to explain what the linear formula looks like by an example. Assume that we project the domain of $x$ and $y$ to the PFA in Figure~\ref{fig:sfa} (a) and (b), respectively. The variables $v_1$, $v_2$, $v_3$, $v_4$ in the figure are \emph{character} variables. One special treatment we do in this paper is that we encoding characters as natural numbers. So $v_1$, $v_2$, $v_3$, $v_4$ are also integer variables. For example, from the formula $ y \in (12)^+$, we may derive $v_3=1$ and $v_4=2$. 
%
%The linear formula produced after the domain restriction will be over variables $v_1$, $v_2$, $v_3$, $v_4$, as well as, the number of occurrences of each character variables $\#v_1$, $\#v_2$, $\#v_3$, $\#v_4$. Each model of the linear formula encodes a model of the string constraint. For example, $x=``12"$ and $y=``1212"$ is encoded by the assignment $(v_1,v_2,v_3,v_4,\#v_1,\#v_2,\#v_3,\#v_4)= (1,2,1,2,1,1,2,2)$. The assignment says, for example, that $x$ is the parametric word obtained by traversing the loop of $A_x$ once (because $\#v_1 = \#v_2 = 1$), which is $v_1v_2$. Under the assignment $v_1=1$ and $v_2=2$, we obtain $x=``12"$.
%
%If a model of the produced linear formula is found, then the procedure concludes SAT with an assignment to the string variables. If not, our procedure step-wisely change the PFA to more expressive ones (by adding more states and transitions) to systematically search for the solution. Depending on the implementation, one may choose to report unknown after some certain steps.
%

@@

%Through the paper, we will use quantifier-free linear integer arithmetic formulae, and call them \emph{linear formulae} for short.  
%Given a linear formula $\phi$, a set of variables $V$, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
%we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$. 
%
%\lh{Our use of indexing by $V$ in the text later is not very nice. Add this for clarity?: If $V$ contain additional variables not present in $\phi$, the these might be assigned arbitrary values by $I$. If $V$ is missing some of the variables of $\phi$, then it must be possible to complete $I$ with assignment of the missing variables into a satisfying assignment of $\phi$ (as in existential quantification).}
%We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
%We drop the reference to $V$ and write just ``model of $\phi$", $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

Through the paper, we will use quantifier-free linear integer arithmetic formulae, and call them \emph{linear formulae} for short.  
Given a linear formula $\phi$ over variables $V$ and an \emph{integer interpretation} of $V$, a function $I:V\rightarrow \integers$, 
we denote by $I\models \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$. 
We use $\modelsof \phi$ to denote the set all models of $\phi$. 
%
%Let $J$ and $J'$ be two sets integer interpretations, $J$ containting interpretation of variables $V$ and $J'$ of variables $V'$. 
%We say that $J$ and $J'$ are \emph{equivalent}, written $J\iequiv J'$, if their interpretations restricted to the common variables are the same, that is, if 
%$\{\restrict I {V\cap V'} \mid I \in J\} = \{\restrict {I'} {V\cap V'} \mid I' \in J'\}$. We may write $J\eqwrt{V\cap V'} J'$ if we wish to made the common variables explicit.

@@

%\section{Under-approximation} \label{section:under_approximate}
%In this section, we describe formally how to restrict the domain of string variables to patterns defined by \emph{parametric flat automata} (PFA) (Section~\ref{section:sfa}) and hence convert string constraints $\insc$ to a set of equisatisfiable linear formulae $\outlf$ in polynomial-time. We assume all disequality constraint $t_1 \neq t2$ are already converted to equal-satisfiable equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}. 
%
%We pick a PFA for each variable in $\vars$ at the beginning of the decision procedure and then construct the set of linear integer constraints in a stepwise manner. In each step, we take a string constraint from $\insc$ and depending on the type of the string constraint, we execute the corresponding procedure to produce the linear integer constraints and add them to the output linear formulae $\outlf$. We also add constraints to $\outlf$ to ensure all character variables of the PFAs only pick values from $\Sigma_\epsilon$.
%
%More concretely, at the beginning, we move all integer constraints in $\insc$ to the output linear formulae $\outlf$. After this step, $\insc$ is the union of the three sets of constraints: the equality constraints, the regular constraints, and the string-number conversion constraints. For all string variables $x$ that occurs in a string-number conversion constraint $n= \sti{x}$, we follow the rule that will be described in Section~\ref{section:s2i} to pick the corresponding PFA. We do not set any restriction to the selection of PFA for other string variables. The selection is an implementation decision. We will describe the PFA selection strategy of \tool in Section~\ref{section:implementation}. 
%
%After the PFA for each variable is selected, we step-wisely pick and remove a string constraint from $\insc$, executing the corresponding procedure to create equisatisfiable linear constraints, and add all the results to $\outlf$. The procedures for regular, equality and string-number conversion constraints are in Sections~\ref{section:mem},~\ref{section:eq}, and~\ref{section:s2i}, respectively. In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.
%
%
% 
%
%
%
%\hide{We replace it with the following constraints.
%	
%	$$
%	\begin{array}{c}
%	(|x_1|+ |x_2|+ \cdots +|x_n| \neq |x_{n+1}| + |x_{n+2}|+ \cdots +|x_m|)\vee \\
%	\left(
%	\begin{array}{cccc}
%	
%	y_1\cdot y_2\cdot y_3 &=& x_1\cdot x_2 \cdots x_n &\wedge\\
%	y_1 \cdot y'_2 \cdot y'_3 &=& x_{n+1}\cdot x_{n+2} \cdots x_m &\wedge\\
%	|y_2|&=&|y'_2|=1
%	\end{array}
%	
%	\right)
%	\end{array}
%	$$
%	
%	The second disjunct says that the constraints $x_1\cdot x_2 \cdots x_n$ and $x_{n+1}\cdot x_{n+2} \cdots x_m$ have a common prefix $y_1$, but the next character $y_2$ and $y'_2$ is different. In an more efficient implementation, one just project $y_2$ and $y'_2$ to a PFA that accept only symbolic words $v_1$ and $v'_1$, respectively (the PFA has no loop and only one transition). }
%
%
%\yfc{say handling regular constraint can be exponential to the number of constraints, but in our under-approximation, it is polynomial.}

@@

\paragraph{Encoding epsilon.}
A minor technical difficulty is that we wish to build flat under-approximation in the form of pure arithmetic formulae, 
but we may also wish to specify when a character variable is assigned $\epsilon$, but $\epsilon$ is not a number. 
Hence, to allow purely arithmetic formulae talk about $\epsilon$, 
we encode $\epsilon$ as some fixed number $\noepsilon\in\nat\setminus\Sigma$.
%In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulae constraining interpretations of character variables.

@@

%$$ % , to the output linear formulae $\phi_O$. 

@@

%Here we give a concrete example. For the equality constraint $xy = yx$, assume that we restrict the domains of $x$ and $y$ to the PFA in Figure~\ref{fig:sfa} (a) and (b), respectively. In the construction of the PFA for $xy$ and $yx$ (Figure~\ref{fig:sfa} (c) and (d)), we rename the variables on the PFA of $X$ and $Y$ to some fresh variable for each different occurrence of $x$ and $y$ in $xy = yx$, connect the PFA after renaming with ``epsilon" transitions (those labeled $v_{\epsilon_0}$ and $v_{\epsilon_1}$), and add linear formulae $v_{\epsilon_0}=-1$ and $v_{\epsilon_1}=-1$ to the output linear formulae $\phi_O$.

@@

%Recall that for flat automata, the number of occurrences of each symbol uniquely characterizes an accepting string (Section~\ref{section:sfa}) and for any finite automaton we can construct a Presburger formula characterizing its Parikh image (Section~\ref{section:preliminary}). For a product automaton $A$ computed from the previous step, we compute a Presburger formula $\phi(A)$ over the set of variables $\{\#{(v_i^{(j)},v_{i'}^{(j')})}\mid (v_i^{(j)},v_{i'}^{(j')}) \in \cvarone \times \cvartwo\}$ characterizing the Parikh image of $A$. We use $\#(v_i)$ and $\#(v_i^{(j)})$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively. The following formulae establish the relation of the number of occurrences of symbols between the PFAs and the product automaton.
